---
title: "Capital Bikeshare"
format: capitalPDF-pdf
params:
  year: 2018
  month: 5
---

```{r}
#| label: load-pkgs
#| echo: false
#| eval: true
#| message: false
library(readr)
library(dplyr)
library(lubridate)
library(showtext)
library(ggplot2)
library(stringr)
```

```{r}
#| label: load-functions
#| echo: false
#| eval: true
#| message: false
source("../functions/plot_circles.R")
source("../functions/plot_map.R")
source("../functions/plot_top5.R")
```

```{r}
#| label: load-fonts
#| echo: false
#| eval: true
#| message: false
font_add_google("Commissioner")
showtext_auto()
```

```{r}
#| label: get-date
#| echo: false
#| eval: true
#| message: false
yr <- as.character(params$year)
mth <- as.character(params$month)
if (nchar(mth) == 1) {
  mth <- paste0("0", mth)
}
filename <- paste0(yr, mth, "-capitalbikeshare-tripdata.csv")
date_str <- paste0(as.character(month(params$month, label = TRUE, abbr = FALSE)), " ", yr)
```

```{r}
#| label: load-data
#| echo: false
#| eval: true
#| message: false
filepath <- paste0("../data/", filename)
OD_data <- read_csv(filepath)
stations <- read_csv("../data/capitalbikeshare-locations.csv")
```

```{r}
#| label: process-data
#| echo: false
#| eval: true
#| message: false
n_journeys <- nrow(OD_data)
members <- table(OD_data$`Member type`)
n_users <- unname(round(100*members["Member"]/sum(members)))
station_list <- unique(c(OD_data$`Start station number`, OD_data$`End station number`))
n_stations <- length(station_list)
station_list <- station_list[which(station_list %in% stations$TERMINAL_NUMBER)]
OD_data <- OD_data |> 
  filter(`Start station number` %in% station_list,
         `End station number` %in% station_list)
```

```{r}
#| label: station-data
#| echo: false
#| eval: true
#| message: false
all_stations <- tibble(station = c(OD_data$`Start station number`, OD_data$`End station number`))
busiest_stations <- all_stations |> 
  group_by(station) |> 
  summarise(n = n()) |> 
  arrange(desc(n)) |> 
  slice_max(n, n = 5) 

top5 <- stations |> 
  select(TERMINAL_NUMBER, ADDRESS) |> 
  inner_join(busiest_stations, by = c("TERMINAL_NUMBER" = "station")) |> 
  arrange(desc(n)) |> 
  pull(ADDRESS)
```

# `r date_str`

Capital Bikeshare is a bicycle-sharing system which serves the Washington, D.C area. In `r date_str`, a total of `r format(n_journeys, big.mark = ",")` trips were made.

## Key figures

This month's summary statistics are shown in @fig-circles.

```{r}
#| label: fig-circles
#| fig-cap: !expr paste0("Key figures for ", date_str) 
#| echo: false
#| eval: true
#| fig-height: 1.5
#| message: false
plot_circles(pickups = n_journeys,
             stations = n_stations,
             users = n_users)
```

## Busiest stations

The busiest station this month is located at `r top5[1]`. The top five busiest stations are highlighted on the map below.

```{r}
#| label: plot-map
#| echo: false
#| eval: true
#| message: false
#| layout-ncol: 2
#| fig-height: 4
plot_map(stations, top5)
plot_top5(top5)
```

### Data

The data included in this report is available from Capital Bikeshare in a public repository at [s3.amazonaws.com/capitalbikeshare-data/index.html](https://s3.amazonaws.com/capitalbikeshare-data/index.html). Further information on any pre-processing performed by Capital Bikeshare is available at [ride.capitalbikeshare.com/system-data](https://ride.capitalbikeshare.com/system-data), and the license agreement at [ride.capitalbikeshare.com/data-license-agreement](https://ride.capitalbikeshare.com/data-license-agreement). 

### Processing

The source code to generate this report can be found at [github.com/nrennie/capital-bikeshare-report](https://github.com/nrennie/capital-bikeshare-report). This report was produced on `r format(Sys.time(), '%b %d, %Y')`.
